---
title: "Storm Data Analysis"
author: "Xiaotao Qu"
date: "February 12, 2016"
output: 
  #html_document:
  knitr:::html_vignette:
    toc: true 
    toc_depth: 2
    number_sections: true
    #theme: journal
    highlight: pygments
---

#Analysis Plan
1. Retreive cBioportal mutation, expression and CNV data for pancreatic cancer
2. stratify patient into two groups, TGFb pathway disrupted, TGFb pathway naive
3. DEG analysis using limma to identify genes that differently expressed in two groups
  3.1 KRAS mutation will be added as a factor


```{r global_options, include=FALSE}
  # use include=FALSE to have the chunk evaluated, but neither the code nor its output displayed.
  knitr::opts_chunk$set(echo=TRUE, message=FALSE, fig.align="center",
                        fig.width=12, fig.height=8, fig.path='figure/',
                        dev='png')
  options(StringsAsFactors=F)
```

```{r echo=F}
  # Clear the current session, to avoid errors from persisting data structures
  rm(list=ls())
  # Free up memory by forcing garbage collection
  invisible(gc())
  # Pretty printing in knitr
  #library(printr)
  # Manually set the seed to an arbitrary number for consistency in reports
  set.seed(1234)
  # Do not convert character vectors to factors unless explicitly indicated
  #options(stringsAsFactors=FALSE)
  startTime <- Sys.time()
```


```{r echo=F}
library(limma)
#library(magrittr)
#library(ggplot2)
library(dplyr)
#library(qvalue)
#library(reshape)
#library(lazyeval)
#library(GGally)
#library(grid)
#library(gridExtra)
library(tidyr)
#library(xtable)
library(printr)
library(rhdf5)
```

#read data in
```{r cache=TRUE}
#system.time(df<-read.csv("data/repdata-data-StormData.csv",stringsAsFactors = F))
#h5createFile("myhdf5file.h5")
#system.time(h5write(df,"myhdf5file.h5","raw"))
#system.time(df<-h5read("myhdf5file.h5","raw"))
#system.time(saveRDS(df,file="raw.Rdata"))
system.time(df<-readRDS(file="raw.Rdata"))
```

#vocabulary control of event type
##controled vocabulary borrowed from National Weather Service Storm Data Documentation
https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf 
page 6 table 2.1.1 Storm Data Event TAble 
given the deadline, manually generated the list to seperate "/" as different event 
add "Landslide" as an alternative for "Debris Flow"
```{r}
data_dict=read.table("Storm_Data_Event_Table.txt",header=T,sep=";")
```

###total number of available event type is `r dim(data_dict)[1]`

##unique entries in the database
```{r}
unique_entries<-unique(df$EVTYPE)
```

###unique entries in the database is `length(unique_entries)`


#generate new data table based on controlled vocabulary
##function that calculate property damage
```{r}
Prop_Damage<-function(base,power_of_ten) {
  #print(paste("number is",base))
  #print(paste("power is",power_of_ten))
  #print(typeof(base))
  if (base == 0) 
    return(0)
  if (power_of_ten=="B") 
    return(base*10^9)
  if (power_of_ten=="K") 
    return(base*10^3)
  if (power_of_ten %in% c("M","m"))
    return(base*10^6)
  return(0)
}
```

```{r}
df.cleaned=df %>% 
    select(EVTYPE,FATALITIES,INJURIES,PROPDMG,PROPDMGEXP,CROPDMG,CROPDMGEXP) %>%
    mutate(damage=Prop_Damage(PROPDMG,PROPDMGEXP)+Prop_Damage(CROPDMG,CROPDMGEXP))

```

```{r cache=TRUE}
df_result=data.frame()
for(event in data_dict$Event.Name) {
  tmp_df=df.cleaned %>% filter(grepl(event,EVTYPE,ignore.case=T)) %>%
    summarise(event=event,
              total_num=n(),
              fatalities=sum(FATALITIES,na.rm=T),
              injuries=sum(INJURIES),
              damage=sum(damage)
    )
  df_result=rbind(df_result,tmp_df)
}
```

