---
title: "Storm Data Analysis"
author: "Xiaotao Qu"
date: "February 12, 2016"
output: 
  #html_document:
  knitr:::html_vignette:
    toc: true 
    toc_depth: 2
    number_sections: true
    #theme: journal
    highlight: pygments
---

```{r global_options, include=FALSE}
  # use include=FALSE to have the chunk evaluated, but neither the code nor its output displayed.
  knitr::opts_chunk$set(echo=TRUE, message=FALSE, fig.align="center",
                        fig.width=12, fig.height=8, fig.path='figure/',
                        dev='png')
  options(StringsAsFactors=F)
```

```{r echo=F}
  # Clear the current session, to avoid errors from persisting data structures
  rm(list=ls())
  # Free up memory by forcing garbage collection
  invisible(gc())
  # Pretty printing in knitr
  #library(printr)
  # Manually set the seed to an arbitrary number for consistency in reports
  set.seed(1203)
  # Do not convert character vectors to factors unless explicitly indicated
  #options(stringsAsFactors=FALSE)
  startTime <- Sys.time()
```


```{r echo=F}
library(limma)
#library(magrittr)
library(ggplot2)
library(dplyr)
library(tidyr)
#library(qvalue)
#library(reshape)
#library(lazyeval)
#library(GGally)
#library(grid)
#library(gridExtra)
library(tidyr)
#library(xtable)
library(printr)
library(rhdf5)
```

#read data in
tried cache and .Rdata to speedy up
```{r reading_data}
#system.time(df<-read.csv("data/repdata-data-StormData.csv",stringsAsFactors = F))
#system.time(saveRDS(df,file="raw.Rdata"))
system.time(df<-readRDS(file="raw.Rdata"))
```

#vocabulary control of event type
##unique entries in the database
```{r}
unique_entries<-unique(df$EVTYPE)
```
unique entries in the database is `r length(unique_entries)` 

these entries are combination of severnal predifined event type 
it makes sense that heavy rain will usually cause flood 

##example of unique event type has "heavy rain" in it
I think they should all combine to one category 
so I try to apply a controlled vocabulary within limited time
```{r}
df %>% 
  select(EVTYPE) %>% 
  filter(grepl("heavy rain",EVTYPE,ignore.case = T)) %>% 
  unique() %>%
  head()
```

##controled vocabulary borrowed from National Weather Service Storm Data Documentation
https://d396qusza40orc.cloudfront.net/repdata%2Fpeer2_doc%2Fpd01016005curr.pdf

page 6 table 2.1.1 Storm Data Event TAble 

given the deadline, manually generated the list to seperate "/" as different event 

add "Landslide" as an alternative for "Debris Flow"
```{r}
data_dict=read.table("Storm_Data_Event_Table.txt",header=T,sep=";")
```

total number of available event type is `r dim(data_dict)[1]`

#generate new data table based on controlled vocabulary
##function that calculate property damage
simple data validation build in as whether the EXP term is meaningful
```{r}
Prop_Damage<-function(base,power_of_ten) {
  #print(paste("number is",base))
  #print(paste("power is",power_of_ten))
  #print(typeof(base))
  if (base == 0) 
    return(0)
  if (power_of_ten=="B") 
    return(base*10^9)
  if (power_of_ten=="K") 
    return(base*10^3)
  if (power_of_ten %in% c("M","m"))
    return(base*10^6)
  return(0)
}
```

##select field of interest and calculate damage as combined property and crops
```{r}
df.cleaned=df %>% 
    select(EVTYPE,FATALITIES,INJURIES,PROPDMG,PROPDMGEXP,CROPDMG,CROPDMGEXP) %>%
    mutate(damage=Prop_Damage(PROPDMG,PROPDMGEXP)+Prop_Damage(CROPDMG,CROPDMGEXP))

```

##generate new data.frame with aggregated value
```{r aggregate_event}
df_result=data.frame()
for(event in data_dict$Event.Name) {
  tmp_df=df.cleaned %>% filter(grepl(event,EVTYPE,ignore.case=T)) %>%
    summarise(event=event,
              total_num=n(),
              fatalities=sum(FATALITIES,na.rm=T),
              injuries=sum(INJURIES),
              damage=sum(damage)
    )
  df_result=rbind(df_result,tmp_df)
}
knitr::kable(head(df_result))
```

#Results
##identify data type with NA value
```{r}
idx_na_row_col=which(is.na(df_result),arr.ind=TRUE)
idx_row=unique(idx_na_row_col[,1])
knitr::kable(df_result[idx_row,])
```

##Which types of events are most harmful with respeiot to population health?
Use death and/or injury per 100 event as indicator of population health for this project 
```{r}
df_result<-df_result %>%
  mutate(death_per_hundred=fatalities/total_num*100,
        injury_per_hundred=injuries/total_num*100
  )

#order data by total heath impact
position=arrange(na.omit(df_result),-(death_per_hundred+injury_per_hundred))$event

#slim the data for plot
df_plot=na.omit(df_result) %>% 
  gather(damage_type,damage_per_hundred,death_per_hundred:injury_per_hundred)
  

ggplot(df_plot, aes(event,damage_per_hundred,fill=damage_type)) + 
  geom_bar(stat="identity")+
  scale_x_discrete(limits=position) +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```

